### 2014-10-31 implementation.tex
edits

@@

>>>>

\subsection{Operation Consistency}

@@

\subsection{Transactions}

While cassandra offers all-or-nothing failure semantics for multiple writes
through batching, readers may witness the initial write while the batch is in
progress. \name implements atomic visibility by exploiting shim layer causality
guarantee -- an effect is included only if all the effects if depends on are
also included.

For every transaction, we instantiate a special transaction marker effect $m$,
but importantly, do not insert into the backing store. $m$ is included as a
dependence to every effect (let $e$ be one such effect) generated in the
transaction. Since $m$ has not yet been written to the store, no operation will
witness $e$ while the transaction in progress. Let $S$ be the set of effects
generated in the transaction. At the end of the transaction, we insert $m$ into
the backing store with $S$ as its dependence. Now any replica which includes
$e$, must include $m$, and transitively must include every effect from the
transaction. This ensures atomicity and satisfies RC requirement.

The above scheme prevents a transaction from witnessing its own effects. This
might conflict with the causality requirement on the operations. Hence,
transactions piggy-back the previous effects from the same transaction for each
request. MAV semantics is implemented by keeping track of the set of
transaction markers $M$ witnessed by the transaction, and before performing an
operation at some replica, ensuring that $M$ is a subset of the transaction
markers included at that replica. If not, the missing effects are synchronously
fetched. RR semantics is realized by capturing a optimized snapshot of the
state of some replica; each operation from an RR transaction is applied to this
snapshot state. Any generated effects are added to this snapshot.

\subsection{Summarization}

The main challenge in realizing an efficient implementation of operation-based
replicated data types is that the state of the object i.e., the set of effects
grows with every effectful operation on the object. If left unchecked, the
operations slow down over time, until the shim layer memory or backing store
disk runs out of memory. Luckily, the state of the operation-based replicated
data type can often be summarized to an \emph{observably equivalent} smaller
state. For example,

\begin{itemize}
\setlength{\itemsep}{2pt}
\item A last-writer-wins register with multiple updates where $v$ is the value
of the last write is observably equivalent to a register with a single write
$v$.

\item A bank account with a series of deposits and withdraws with current
balance $b$ is equivalent to a bank account with a single deposit of $b$.

\item A set with collection of add and remove operations is equivalent to a set
with a series of add operations of live elements from the original set.
\end{itemize}

Since the semantics of summarization depends on the semantics of the data type,
we expect the programmer to provide a summarization function for each RDT with
the following type:

\begin{codehaskell}
summarize :: Effect e => [e] -> [e]
\end{codehaskell}

\noindent with the intention that the length of the result is smaller that the
length of the argument. We utilize the \cf{summarize} function to summarize the
object state both in the shim layer node and the backing store, typically when
the number of effects on an object crosses a tunable threshold. Shim layer
summarization is straight-forward; a summarization thread takes the local lock
on the object, and replaces its state with the summarized state. The shim layer
node remains unavailable for that particular object during summarization
(usually a few milliseconds).

Compared to the shim layer, summarization in the backing store is more
complicated. The main challenge is that unlike the shim layer, summarization
cannot run as an atomic operation. Summarization in the backing store involves
deleting previously inserted rows and inserting new rows, where each row
corresponds to an effect. It is essential that concurrent client operations are
permitted, but are not allowed to witness the intermediate state of the
summarization process.

To this end, we adopt a novel summarization strategy that builds on the
causality property of the store. Suppose we have a set $O$ of effects on a
object, which when summarized yields a new set of effects $N$. We first
instantiate a summarization marker $s$, and similar to transaction marker, we
do not insert it into the store immediately. We insert the new effects in $N$
with strong consistency, including $s$ as a dependence. Since $s$ is not yet in
the store, effects from $N$ are not made visible to the clients. Then we insert
$s$ with strong consistency, including $O$ as dependence, and finally delete
the rows corresponding to $O$. When the shim layer node witnesses the
summarization marker $s$ on the object, it removes the effects from $O$ from
its cache. But due to strong consistency of insertion of $N$ and $s$, the new
effects $N$ are also observed. Thus, the clients witness the observably
equivalent new set of effects $N$, and not the old effects $O$. This ensures
atomicity of summarization in the backing store.
